---
title: "Exercises 02"
author: "Bjarne Taulo Sørensen"
date: "Week 5 and 6 - 2022"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br />\    
  
  
<!-- Your own secret notes can be written here between the HTML signs -->   



```{r bibliotek, eval=TRUE, include=TRUE}

if (!dir.exists("data")) dir.create("data")

# Gem dine datasæt i dette bibliotek

```


```{r pacman, eval=TRUE, include=TRUE}

pacman::p_load("tidyverse", "magrittr", "nycflights13", "gapminder",
               "Lahman", "maps", "lubridate", "pryr", "hms", "hexbin",
               "feather", "htmlwidgets", "broom", "pander", "modelr",
               "XML", "httr", "jsonlite", "lubridate", "microbenchmark",
               "splines", "ISLR2", "MASS", "testthat", "leaps", "caret",
               "RSQLite", "class", "babynames", "nasaweather",
               "fueleconomy", "viridis", "boot", "rvest", "stringi", 
               "readxl", "haven")

```



<br />\   

<!-- Your own secret notes can be written here between the HTML signs -->


# Problem 1 - Do not solve this one

Get this data:

```{r problem 1, eval=FALSE, include=TRUE}

my_url01 <- "http://tiny.cc/dcf/houses-for-sale.csv"

```

Are the variables in the right format?

```{r problem 1 solution, eval=FALSE, include=FALSE}

my_url01 <- "http://tiny.cc/dcf/houses-for-sale.csv"
houses <- readr::read_csv(my_url01) 
str(houses) # Yes

# devtools::install_github("ekstroem/dataReporter")
library(dataReporter)

makeDataReport(data = houses)


```

Remember to make comments.


# Problem 2

Now, get these data:

```{r problem 2, eval=FALSE, include=TRUE}

my_url02 <- "https://perso.telecom-paristech.fr/eagan/class/igr204/data/factbook.csv"

```
Remember to make comments.

```{r problem 2 solution, eval=FALSE, include=FALSE}

countries01 <- readr::read_csv2(my_url02, trim_ws = TRUE) # Does this code work on your computer?

countries02 <- read_delim(my_url02,";", escape_double = FALSE, 
                        locale = locale(decimal_mark = "."), 
                         trim_ws = TRUE) # Is this working?

```

The second line is not needed, i.e. where Country = "String", so please, delete it. "Skip = 1"

Make sure that the variables are in a proper format! 

Anything strange about the names of the variables?



```{r problem 2 solution a, eval=FALSE, include=FALSE}

# saveRDS(countries01, file = "data/countries01.Rda")

countries01 <- readRDS(file = "data/countries01.Rda")

#1 if data file is located on the internet:
str(countries02)

?read_csv2
headers <- read_csv2(my_url02, col_names = FALSE, n_max = 1)
countries03 <- read_delim(my_url02, delim = ";", skip = 2, col_names = FALSE) 
colnames(countries03) = headers

str(countries03)

# if data file from Moodle in RDA format:

countries_tidyverse <- readRDS(file = "data/countries01.Rda") %>%
  filter(Country != "String") %>% 
   mutate(
  `Area(sq km)` = as.double(`Area(sq km)`),
  `Birth rate(births/1000 population)` = as.double(`Birth rate(births/1000 population)`))

countries_tidyverse <- readRDS(file = "data/countries01.Rda") %>%
  filter(Country != "String") %>% 
   mutate(
  `Area(sq km)` = as.double(`Area(sq km)`),
  `Birth rate(births/1000 population)` = as.double(`Birth rate(births/1000 population)`))
  #ETC - can do this for all variables that we want to parse to a double format.
  #You can also use parse_double instead of as.double
  #We are going to make this smarter later on - map and apply functions - looping over the elements of the data frame. 
```

# Problem 3

Remember to make comments.

Get (read) all the tables from: 

```{r problem 3, eval=FALSE, include=TRUE}
my_url03 <- "http://en.wikipedia.org/wiki/Mile_run_world_record_progression"

```

Get help from *Google*!  

*Hint*: **read_html** and **html_nodes("table")**  

Let's have a look at table 2.

```{r problem 3 solution, eval=FALSE, include=FALSE}
my_url03 <- "http://en.wikipedia.org/wiki/Mile_run_world_record_progression"
?readr

tables01 <- my_url03 %>% 
  read_html() %>% 
    html_nodes("table") # We end up with a list

length(tables01)

tables01a <- html_table(tables01[[2]]) %>% 
  View()
```


# Problem 4


Remember to make comments.
  
Here you can find all-time top-25 marathoners - women/men:  
  
```{r problem 4, eval=TRUE, include=TRUE}

my_url04 <- "https://en.wikipedia.org/wiki/Marathon"

```
Please make two data sets, one for women and one for men showing top-25 marathoners:

```{r problem 4 solution, eval=FALSE, include=FALSE}

tables02 <- my_url04 %>% 
  read_html() %>% 
   html_nodes("table")

View(tables02)
length(tables02)

marathon_women <- html_table(tables02[[6]], fill = TRUE) %>% 
  View()
marathon_men <- html_table(tables02[[5]], fill = TRUE) %>% 
  View() 
```

# Problem 5

Remember to make comments.

Import **file01**, **file02**, **file03**, **file04**, **file05**, and **file06**

```{r problem 5 solution, eval=FALSE, include=FALSE}

file01 <- read_csv("data/file01.csv")
str(file01)
                   
file01 <- read_csv("data/file01.csv", col_types = cols(
  DOB = col_date(format = "%Y-%m-%d"), 
   Died = col_date(format = "%Y-%m-%d")), 
                   locale = locale()
                                                           )
str(file01)  


file02 <- read_delim("data/file02.csv", ";", escape_double = FALSE, 
                     trim_ws = TRUE)


file03 <- read_delim("data/file03.txt", "|", escape_double = FALSE, 
                     col_types = cols(
                       DOB = col_date(format = "%Y-%m-%d"), 
                       Died = col_date(format = "%Y-%m-%d")
                                    ), 
                     trim_ws = TRUE)


file04 <- read_table2("data/file04.fwf", 
                     col_types = cols(
                       DOB = col_date(format = "%Y-%m-%d"), 
                       Died = col_date(format = "%Y-%m-%d")
                                      )
                     )


file04 <- read_table("data/file04.fwf", 
                     col_types = cols(
                       DOB = col_date(format = "%Y-%m-%d"), 
                       Died = col_date(format = "%Y-%m-%d")
                                      )
                     )

# Fixed-width typeface.


file05 <- read_delim("data/file05.tsv", "\t", 
                    escape_double = FALSE, 
                     col_types = cols(
                      DOB = col_date(format = "%Y-%m-%d"), 
                      Died = col_date(format = "%Y-%m-%d")
                                     ), 
                     trim_ws = TRUE)


file06 <- read_delim("data/file06.csv", ";", escape_double = FALSE, trim_ws = TRUE)

file06 <- read_delim("data/file06.csv", ";", escape_double = FALSE,
                     col_types = cols(
                       Name = col_character(),
                        DOB = col_date(format = "%Ybeep%mbeep%d"),
                         Birthplace = col_character(),
                          Died = col_date(format = "%Ynotahyphen%mnotahyphen%d"),
                           Occupation = col_character(),
                            Children = col_double()
                                    ),
                     trim_ws = TRUE)


```

# Problem 6

Remember to make comments.

Import the following data:

```{r problem 6, eval=FALSE, include=TRUE}
my_url05 <- "https://perso.telecom-paristech.fr/eagan/class/igr204/data/nat1900-2017.tsv"

names_fr <- readr::read_delim(my_url05, "\t")

table(names_fr$sexe)
distinct(names_fr, sexe)
# table(names_fr$preusuel)
# table(names_fr$annais)
# table(names_fr$nombre)
```



Keep only females, i.e. **sexe** should be *2* . 

Drop observations where **preusuel** is _PRENOMS_RARES.

Some rare names don't exist some of the years. Make sure that, if a name is not registered a given year, then **nombre** should be 0, and not missing. (Hint: first use **complete()**, then use **ifelse()** in connection with **mutate()**).

**sexe** should be *2* instead of missing. 

Delete an observation if the name is missing. 

Drop the variable **sexe**. 

Delete observations if **annais** is equal to **XXXX**.

```{r problem 6 solution a, eval=FALSE, include=FALSE}
names_fr <- readr::read_delim(my_url05, "\t") %>%
  filter(sexe == 2, !is.na(preusuel), 
         annais != "XXXX", preusuel !="_PRENOMS_RARES") %>% 
   complete(preusuel, annais) %>% 
    mutate(nombre = ifelse(is.na(nombre), 0, nombre), 
           sexe = ifelse(is.na(sexe), 2, sexe)) %>% 
     dplyr::select(-sexe)

view(names_fr)
```

Which female name has the highest frequency - a given year and all time?

Keep the variables **preusuel**, **annais**, **nombre**

```{r problem 6 solution b, eval=FALSE, include=FALSE}

max_pr_year <- names_fr %>% 
  group_by(annais) %>% 
   filter(nombre == max(nombre)) %>% 
     arrange(annais) %>% 
      ungroup() %>% 
       dplyr::select(preusuel, annais, nombre) # the only variables we got

max_all_time <- max_pr_year %>% 
  filter(nombre == max(nombre))


```

# Problem 7

Show the number of people named **Marie** from 1900 to 2017 in a **plot**. 

Make some x and y **labels**, and a **title**. 

Moreover, add a **theme**.

```{r problem 7, eval=FALSE, include=FALSE}
names_fr_marie <- names_fr %>% 
  filter(preusuel == "MARIE") 

ggplot2::ggplot(data = names_fr_marie, mapping = aes(x = annais, y = nombre)) +
 geom_point() +
  scale_x_discrete(breaks=seq(1900, 2017, 10)) +
   ggtitle("Marie for more than 100 years") +
    xlab("Years") + 
     ylab("Count") +
      theme_bw() 
      
```

# Problem 8

Import the SPSS file: **survey.sav**, the SAS file: **x2012_1_6.sas7bdat**, and the Excel file: **excel_file**. 

```{r problem 8, eval=FALSE, include=FALSE}

survey <- read_sav("data/survey.sav")

x2012_1_6a <- read_sas("data/x2012_1_6.sas7bdat", NULL)


excel_file <- read_excel("data/excel_file.xlsx", 
    col_names = FALSE)
excel_file <- read_excel("data/excel_file.xlsx", 
    sheet = "Number02", skip = 10)
excel_file <- read_excel("data/excel_file.xlsx", 
    sheet = "Number03", range = "E6:G9")

```



# Problem 9

The following code provides ggplot of sumprice and weekday. **søndag** = Sunday is listed before **lørdag** = Saturday. I want the weekdays in the following order:
**søndag = Sunday, lørdag = Saturday, fredag = Friday, torsdag = Thursday, onsdag = Wednesday, tirsdag = Tuesday, mandag = Monday**

```{r problem 9, eval=TRUE, include=TRUE}

Canteen_clean <- readRDS(file="data/Canteen_clean.Rda")

ggplot(data = Canteen_clean) + 
 geom_boxplot(mapping = aes(x = reorder(Weekday, sumprice, FUN = median), y = sumprice)) + 
  coord_flip()


```

Make some code that will correct the order of the weekdays. And translate into English.

Remember to make comments.

```{r problem 9 solution, eval=FALSE, include=FALSE}
class(Canteen_clean$Weekday)

Weekday_levels <- c("mandag", "tirsdag", "onsdag", "torsdag", "fredag", "lørdag", "søndag")

Canteen_clean_factor <- Canteen_clean %>% 
  mutate(WeekdayF = factor(Weekday, levels = Weekday_levels))  %>%  
   dplyr::select(WeekdayF, everything())  %>%
    mutate(WeekdayF = fct_recode(WeekdayF,
                               "Monday" = "mandag",
                               "Tuesday" = "tirsdag",
                               "Wednesday" = "onsdag",
                               "Thursday" = "torsdag",
                               "Friday" = "fredag",
                               "Saturday" = "lørdag",
                               "Sunday" = "søndag"
  ))

class(Canteen_clean_factor$WeekdayF)


ggplot(data = Canteen_clean_factor) +
  geom_boxplot(aes(WeekdayF, sumprice)) + 
   coord_flip()



```



# Problem 10

Redo the following code using the pipe **%>%** Make only one new data set.

```{r problem 10, eval=FALSE, include=TRUE}

popu_dest01 <- flights
popu_dest02 <- group_by(popu_dest01, dest)
popu_dest03 <- mutate(popu_dest02, n=n())
popu_dest04 <- dplyr::select(popu_dest03, n, everything())
popu_dest05 <- filter(popu_dest04, n > 10000)


```

```{r problem 10 solution, eval=FALSE, include=FALSE}

one_new_data_set01 <- flights %>% 
 group_by(dest) %>% 
  mutate(n=n()) %>% 
   dplyr::select(n, everything()) %>% 
    filter(n > 10000)

compare(popu_dest05, one_new_data_set01)

```

# Problem 11

Redo the following code using the pipe **%>%** Make only one new data set.

```{r problem 11, eval=FALSE, include=TRUE}

one_new_data_set02 <- filter(
                       dplyr::select(
                        mutate(
                         group_by(flights, dest), 
                               n=n()), 
                              n, everything()), 
                             n > 10000)

```

```{r problem 11 solution, eval=FALSE, include=FALSE}

one_new_data_set02 <- filter(dplyr::select(mutate(group_by(flights, dest), n=n()), n, everything()), n > 10000)

one_new_data_set02 <- flights %>% 
 group_by(dest) %>%
  mutate(n=n()) %>% 
   dplyr::select(n, everything()) %>% 
    filter(n > 10000)

  
compare(one_new_data_set01, one_new_data_set02)

```




# Problem 12

Based on **yq**, tidy up. Sort the data into chronological order.


```{r problem 12, eval=FALSE, include=TRUE}
yq <- tribble(
~group, ~y, ~q1, ~q2, ~q3, ~q4,
"a1",	2016,	25,	36,	30,	47,
"a1",	2017,	21,	33,	37,	43,
"a1",	2018,	22,	32,	34,	40,
"a1",	2019,	20,	34,	30,	46,
"a2",	2016,	22,	33,	30,	40,
"a2",	2017,	26,	34,	31,	40,
"a2",	2018,	23,	31,	30,	45,
"a2",	2019,	23,	30,	36,	40,
"a3",	2016,	20,	32,	32,	46,
"a3",	2017,	23,	31,	37,	41,
"a3",	2018,	27,	32,	33,	40,
"a3",	2019,	24,	30,	31,	44)
```

```{r problem 12 solution, eval=FALSE, include=FALSE}

New_yq <- yq %>% 
  pivot_longer(cols = q1:q4,  names_to = "quarter", values_to = "sales") %>%
   arrange(group, y, quarter)
view(New_yq )
```

Compare your result to:

```{r problem 12a , eval=FALSE, include=TRUE}

yq %>% pivot_longer(cols = -c(group, y), names_to = "quarter", values_to = "sales") %>% arrange(y, quarter)
yq %>% pivot_longer(cols = 3:6, names_to = "quarter", values_to = "sales") %>% arrange(y, quarter)

```




# Problem 13

```{r problem 13 , eval=FALSE, include=TRUE}
Canteen_data_inRformat <- readRDS(file="data/data20171001.Rda")
Extra001 <- Canteen_data_inRformat %>% 
  filter(!trimws(Department) %in% c(NA, "-", "")) %>% 
  dplyr::select(-Segment, -Units01) %>% 
  mutate(id = row_number())
Extra002 <- Extra001 %>% 
  filter(!(!Subsidy %in% c(" ") & Price01 < 0)) %>% 
  filter(Price01 > 0) 
Extra003 <- Extra001 %>% 
  filter(!Subsidy %in% c(" ")) %>% 
  mutate(id_extra = as.integer(id - 1)) %>% 
  dplyr::select(-Date01, -Timofday03, -(Department:Product), -id) %>% 
  rename(Initials_extra = Initials, Sub_extra = Subsidy, Price_extra = Price01)


```


Merge Extra002 and Extra003. Use SQL!

Remember to make comments.  

```{r problem 13 solution, eval=FALSE, include=FALSE}

if (!dir.exists("ourdata")) dir.create("ourdata")

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")

dbWriteTable(connection_ourdata, "Extra001", Extra001, row.names = FALSE, overwrite = TRUE)
dbWriteTable(connection_ourdata, "Extra002", Extra002, row.names = FALSE, overwrite = TRUE)
dbWriteTable(connection_ourdata, "Extra003", Extra003, row.names = FALSE, overwrite = TRUE)
dbDisconnect(connection_ourdata)


connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
dbListFields(connection_ourdata, "Extra001")
dbListFields(connection_ourdata, "Extra002")
dbListFields(connection_ourdata, "Extra003")
dbDisconnect(connection_ourdata)

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
EXTRA2and3_db  <- dbGetQuery(connection_ourdata, 
                             "select s.*, t.*
                                from Extra002 as s 
                                left join Extra003 as t 
                                on s.id = t.id_extra
                             ")
dbDisconnect(connection_ourdata)

# View(EXTRA2and3_db)   


#Extra
connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
db_Extra001 <- dbReadTable(connection_ourdata, "Extra001")
db_Extra002 <- dbReadTable(connection_ourdata, "Extra002")
db_Extra003 <- dbReadTable(connection_ourdata, "Extra003")
dbDisconnect(connection_ourdata)

```

# Problem 14

Based on Extra001, keep observations having “product” = “Pant retur” (i.e. bottle deposits). Make a summary for each employee – i.e. how many observations per employee, and how much money is paid back to 
the employees in total. Keep only observations if the sum of Price01 is less than -10. Use SQL!  

Remember to make comments.  

```{r problem 14 solution, eval=FALSE, include=FALSE}

connection_ourdata <- dbConnect(SQLite(), "ourdata/r_db_ex.sqlite")
EXTRA_deposit_db <- dbGetQuery(connection_ourdata, 
                               "select *, count(*) as no, sum(Price01) as pr
                                from Extra001
                                 where trim(product) = 'Pant retur'
                                  group by Initials
                                   having pr < -10"
                               )
dbDisconnect(connection_ourdata)
```
