---
title: "Exercises 03"
author: "Bjarne Taulo Sørensen"
date: "Week 5 & 6 - 2022"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br />\    
  
  
<!-- Your own secret notes can be written here between the HTML signs -->   



```{r bibliotek, eval=TRUE, include=TRUE}

if (!dir.exists("data")) dir.create("data")

# Gem dine datasæt i dette bibliotek

```


```{r pacman, eval=TRUE, include=TRUE}

# devtools::install_github("ekstroem/dataReporter")
pacman::p_load("tidyverse", "magrittr", "nycflights13", "gapminder",
              "Lahman", "maps", "lubridate", "pryr", "hms", "hexbin",
              "feather", "htmlwidgets", "broom", "pander", "modelr",
              "XML", "httr", "jsonlite", "lubridate", "microbenchmark",
              "splines", "ISLR2", "MASS", "testthat", "leaps", "caret",
              "RSQLite", "class", "babynames", "nasaweather",
              "fueleconomy", "viridis", "boot", "devtools", "rvest",
              "stringi", "readxl", "haven", "dataReporter")
```



<br />\   

<!-- Your own secret notes can be written here between the HTML signs -->


# Problem 1


Make a random sample from **X201204** of 500 distinct employees. Make sure that these are not from *Afdeling=”AgroTech A/S”* and not part of the **sample01** data set.  

```{r problem 1 solution, eval=FALSE, include=FALSE}

sample01 <- readRDS(file="data/sample01.Rda")

x201204a <- readRDS(file="data/x201204.Rda") %>% 
  filter(!Afdeling == 'AgroTech A/S') %>% 
   filter(!Initialer %in% sample01$Initials) %>%
    filter(!trimws(Initialer) %in% c(NA, "-", "")) %>%
     dplyr::select(Initialer) %>% 
      distinct()

# Solution 1
set.seed(23)
x201204b <- x201204a %>% 
 mutate(xxx = sample(nrow(x201204a), size = nrow(x201204a), replace = FALSE)) %>% 
  arrange(xxx) %>% 
   head(500) 
nrow(x201204b)

# Solution 2 easier
set.seed(911)
x201204c <- x201204a %>% 
 sample_n(50)
nrow(x201204c)

```


# Problem 2

Use SQL!  

Use the **flights** data and:  
  
1)	**airports**, *merge* by *origin=faa*

a.	keep the following variables: *year*, *month*, *day*, *carrier*, *flight*, *tailnum*, *origin*, *dest*, *distance*, *name*
b.	condition: output if *distance* greater than 2.5 the average *distance*.

2)	**planes**, *merge* i.e.

c.	*merge* the total sum of *flight* for each *tailnum* from **flights** with **planes**, but only if the total sum of *flight* for a given *tailnum* is greater than 10000.

3)	**weather**

d.	select all variables from **flights**, but keep only observations where *origin* from **weather** starts with an E and has two additional letters.

4)	**airlines**

e.	select all variables from **flights**, but drop observations where *name* from **airlines** includes lines or Lines.

Remember to make comments.  

```{r problem 2 solution, eval=FALSE, include=FALSE}

if (!dir.exists("ourdata03")) dir.create("ourdata03")
con <- dbConnect(SQLite(), "ourdata03/flights_db.sqlite")

dbWriteTable(con, "airports", airports, row.names = FALSE, overwrite = TRUE)
dbWriteTable(con, "planes", planes, row.names = FALSE, overwrite = TRUE)
dbWriteTable(con, "flights", flights, row.names = FALSE, overwrite = TRUE)
dbWriteTable(con, "weather", weather, row.names = FALSE, overwrite = TRUE)
dbWriteTable(con, "airlines", airlines, row.names = FALSE, overwrite = TRUE)

dbListTables(con)

# 1

flights_airports <- dbGetQuery(con, 
                               "select s.year, s.month, s.day, s.carrier, s.flight, 
                                 s.tailnum, s.origin, s.dest, s.distance, t.name
                                  from flights as s
                                   left join 
                                    airports as t
                                     on s.origin = t.faa
                                      where s.distance > (select 2.5*avg(distance) from flights)
                               ")

# 2

flights_planes <- dbGetQuery(con,
                             "select s.*, t.* from 
                               (select tailnum, sum(flight) as sumflight from flights 
                                group by tailnum
                                 having sumflight > 10000) as s
                                  left join 
                                   planes as t
                                    on s.tailnum=t.tailnum
                             ")

# 3

flights_weather <- dbGetQuery(con,
                              "select *
                                from flights
                                 where trim(origin) in (select distinct origin 
                                  from weather where trim(origin) like 'E__')
                              ")

# 4

flights_airlines <- dbGetQuery(con,
                               "select *
                                 from flights
                                  where trim(carrier) not in (select distinct carrier 
                                   from airlines where trim(name) like '%lines%' or 
                                    trim(name) like '%Lines%')
                               ")
```



# Problem 3

Make comments to the following code:

```{r problem 3, eval=FALSE, include=TRUE}

Hide <- function(x){
	cat("You entered the following number:", x)
 	invisible(x^1 + x^2 + x^3 + x^4 + x^5)
 }

Hide(2)
y <- Hide(2)
y

```

# Problem 4

Make comments to the following code:

```{r problem 4, eval=FALSE, include=TRUE}

set.seed(123)
y <- stats::rnorm(10000000)
table(cut(y, breaks = -6:6))

```



# Problem 5

What is the difference: 

```{r problem 5, eval=FALSE, include=TRUE}

# I

show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  invisible(df)
}

mtcars %>% 
  show_missings() %>%  
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings() 


# II

show_missings_II <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
}

mtcars %>% 
  show_missings_II() %>%  
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings_II() 



```


# Problem 6

Use *strsplit* to split the strings and make each part accessible. Moreover, make the data tidy.
Finally, make sure you end up with a *tibble*. Again, remember to make comments.

```{r problem 6, eval=FALSE, include=TRUE}

car_french <- c("Product, Peugeot 108, Quantity, 2000, Year, 2019",
                "Product, Peugeot 108, Price, 20000, Year, 2019",
                "Product, Peugeot 208, Quantity, 1810, Year, 2018",
                "Product, Peugeot 208, Price, 23400, Year, 2018",
                "Product, Peugeot 308, Quantity, 1005, Year, New",
                "Product, Peugeot 308, Price, 32500, Year, New"
                ) 

```

```{r problem 6 solution, eval=FALSE, include=FALSE}

car_french01 <- as_tibble(do.call(rbind, strsplit(car_french, ", ", fixed=TRUE))) %>% 
  pivot_wider(names_from = V3, values_from = V4 ) %>% 
    rename(Product = V2, Year = V6) %>% 
      dplyr::select(Product, Year, Price, Quantity) %>% 
       mutate(Year = parse_double(ifelse(trimws(Year) == "New", 2020, Year)), 
        Price = parse_double(Price), Quantity = parse_double(Quantity))  

car_french01

str(car_french01)

```


# Problem 7

Please make a new similar program without using SQL. Assume that the data set **diamonds**
exists outside the data base. Again, remember to make comments.

```{r problem 7, eval=FALSE, include=TRUE}

db_diamonds <- dbGetQuery(con,
 "select *, price / size as value_density from
 (select carat, price, x * y * z as size from diamonds)"
 )

head(db_diamonds, 3)

```

```{r problem 8 solution, eval=FALSE, include=FALSE}

diamonds01 <- diamonds %>% 
 mutate(size = x*y*z, value_density = price/size) %>% 
  dplyr::select(carat, price, size, value_density) %>% 
   head(3)

```

# Problem 9

Use the functions **semi_join()** and **anti_join()** instead of the first two **filter()** functions.

```{r problem 9, eval=FALSE, include=TRUE}

sample01 <- readRDS(file="data/sample01.Rda")
sample02 <- readRDS(file="data/sample02.Rda")

x1 <- readRDS(file="data/x201205.Rda") %>% 
 filter(Initialer %in% sample02$Initials) %>%
  filter(!Initialer %in% sample01$Initials) %>%
   filter(!trimws(Initialer) %in% c(NA, "-", "")) %>%
    dplyr::select(Initialer) %>% 
     distinct()

nrow(x1)


```

```{r problem 9 solution, eval=FALSE, include=FALSE}

sample01 <- readRDS(file="data/sample01.Rda")
sample02 <- readRDS(file="data/sample02.Rda")

x2 <- readRDS(file="data/x201205.Rda") %>% 
 semi_join(sample02, by = c("Initialer"="Initials")) %>% 
  anti_join(sample01, by = c("Initialer"="Initials")) %>% 
   filter(!trimws(Initialer) %in% c(NA, "-", "")) %>%
    dplyr::select(Initialer) %>% 
     distinct()

nrow(x2)

identical(x1, x2)

```

Remember to make comments. 

# Problem 10

Append **a** and **b**, **a** and **c**, and **a** and **d**:

```{r problem 10, eval=FALSE, include=TRUE}

a <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3",
  4, "x4"
)

b <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2"
)

c <- tribble(
  ~key, ~val_x,
  1, "z1",
  2, "z2",
  3, "z3",
  4, "z4"
)

d <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3",
  4, "x4"
)

```

Remember to make comments. 
```{r, problem 10 solution, eval=FALSE, include=FALSE}

dplyr::union_all(a, b)
dplyr::union_all(a, c)
dplyr::union_all(a, d)
  

dplyr::union(a, b) # Not!
dplyr::union(a, c)
dplyr::union(a, d)
```


# Problem 11

Please make a new similar program without using the *switch()* function. Again, remember to make
comments.

```{r problem 11, eval=FALSE, include=TRUE}

MMSD <- function(y, tp) {
                    switch(tp, mean = mean(y), 
                               median = median(y), 
                               sd = sd(y)
                           )
                         }  
y<-c(1, 4, 3, 7, 5, 9, 7, 3, 9, 1, 12)

MMSD(y, "mean")
MMSD(tp="mean", y=y)
MMSD(y, "MEDIAN")

```

```{r problem 11 solution, eval=FALSE, include=FALSE}

MMSD2 <- function(y, tp) {
          if (tp == "mean") {
             mean(y)
          } else if (tp == "median") {
             median(y)
          } else if (tp == "sd")  {
             sd(y)
          } #else {
            # stop("Unknown type of operation")   
         # }
}

y<-c(1, 4, 3, 7, 5, 9, 7, 3, 9, 1, 12)
  
MMSD2(y, "mean")
MMSD2(y, "MEDIAN")

```


# Problem 12

Make comments to the following code:

```{r problem 12, eval=FALSE, include=TRUE}

bts <- c("XXXXXXbla-19960101T000000Z-1.tsv",
         "XXXXXXbla-19960101T000000Z-2.tsv", "XXXXXXbla-19960101T000000Z-3.tsv")

 Datafr <- data.frame(do.call(rbind, strsplit(bts, "-", fixed=TRUE)))
  Datafr[,2] <- as.Date(Datafr[,2] , format="%Y%m%d")
   Datafr[,3] <- as.integer(gsub(".tsv", "", Datafr[,3], fixed=TRUE))

```


# Problem 13

The data set **Exam** needs some cleaning up. First, remove observations if **idnumber02** is missing unless the employee is a woman, i.e. **Male=2**. Next, remove an observation if the employee is older than *68* (**Age**), or **Initials** contain a number. Call this new data set for **Exam3**.

Remember to make comments.  

```{r problem 13 solution, eval=FALSE, include=FALSE}
exam <- readRDS(file="data/exam.Rda")
    
exam_a <- exam[complete.cases(exam[ , 5]),]

exam_b <- exam %>% 
 filter(Male==2)

Exam3 <- rbind(exam_a, exam_b) %>% 
   distinct() %>% 
    filter(Age < 68) %>%  # accept 69 also 
     filter(!str_detect(initials, "[[:digit:]]"))
View(Exam3)

```

# Problem 14

Make comments to the following code:

```{r problem 14 , eval=FALSE, include=TRUE}

data_vector = c(9, 0.2, 5, 1, 23, 7, 2, 2, 5, .53) 
 categories = cut(data_vector, breaks=c(0,1,5,max(data_vector)), right = TRUE) 
  categories
   class(categories)
    typeof(categories)
     table(categories) 
      levels(categories) = c("Good","Better","Best")
       table(categories)
     
```

# Problem 15

Remember to make comments.  


First, extract the 2nd column from the list of matrices (i.e. from each matrix).

Next, extract the 3rd row from the list of matrices.

Hint: use lapply and [.

```{r problem 15, eval=FALSE, include=TRUE}

x1 <- matrix(1:30, 3)
x2 <- matrix(2:22, 3)
x3 <- matrix(117:143, 3)
x4 <- matrix(66:95, 5)
list_of_matrices = list(x1, x2, x3, x4)
list_of_matrices
```

```{r problem 15 solution, eval=FALSE, include=FALSE}

lapply(list_of_matrices,"[", , 2)
lapply(list_of_matrices,"[", 3, )

?'['
```


# Problem 16

Make comments to the following code:

```{r problem 16, eval=FALSE, include=TRUE}

apply(iris[,1:4], 2, boxplot)


round(apply(ggplot2::diamonds[10000:11000, 8:10], 2, mean), 1)

purrr::flatten_dbl(purrr::map(-2:2, rnorm, n = 6))

?rnorm
?flatten_dbl
purrr::map(-2:2, rnorm, sd = 2, n = 6)
```


# Problem 17

Make a program that rolls a die multiple times. The loop should continue until you get 8 sixes in a row. Calculate the total number of rolls. 

Remember to make comments.  

```{r problem 17 solution, eval=FALSE, include=FALSE}

roll_die <- function() sample(c("One", "Two", "Three", "Four", "Five", "Six"), 1)

total <- 0
n_six <- 0

while (n_six < 8) { 
  if (roll_die() == "Six") { 
    n_six <- n_six + 1
  } else {
    n_six <- 0
  }
  total <- total + 1
}

total

```


# Problem 18

Make a function that checks if a word is uppercase:

Remember to make comments.  

```{r problem 18, eval=FALSE, include=FALSE}

upper <- function(x) {
  str_detect(x, "^[:upper:]+$")
}

rfba <- c("business", "BUSINESS", "BusinesS", "")
upper(rfba)

```


# Problem 19

Make comments to the following code:

```{r problem 19, eval=FALSE, include=TRUE}

string_rfba <- "11 april 2020"
regexp <- "([[:digit:]]{2}) ([[:alpha:]]+) ([[:digit:]]{4})"
x <- sub(regexp, "\\1", string_rfba)
y <- sub(regexp, "\\2", string_rfba) 
z <- sub(regexp, "\\3", string_rfba) 

```


# Problem 20

Remember to make comments.  

```{r problem 20, eval=FALSE, include=TRUE}

mtcars %>%
  split(.$cyl) %>%
  purrr::map(~ lm(mpg ~ wt, data = .)) %>%
  purrr::map(summary) %>%
  purrr::map_dbl("r.squared")

?split
```


Remember to make comments. 

# Problem 21

Make comments to the following code:

```{r problem 21, eval=FALSE, include=TRUE}

geoms_bts = list(
  geom_point(),
  geom_line(),
  geom_smooth())

ggplot(mtcars, aes(hp, mpg)) + geoms_bts

```



# Problem 22

Below, I have created a vector of all the mean values of all the columns of mtcars. Make some new code that uses a for loop to create the means. Can you also use lapply, sapply, purrr::map, map_lgl, map_int, map_dbl, and map_chr?


```{r problem 22, eval=FALSE, include=TRUE}
vectors_of_means <- c(mean(mtcars$mpg), mean(mtcars$cyl), 
                      mean(mtcars$disp), mean(mtcars$hp),
                      mean(mtcars$drat), mean(mtcars$wt), 
                      mean(mtcars$qsec), mean(mtcars$vs),
                      mean(mtcars$am), mean(mtcars$gear), 
                      mean(mtcars$carb))
```


```{r problem 22 solution, eval=FALSE, include=FALSE}
loop_vector_means <- vector("double", ncol(mtcars))

for (i in seq_along(mtcars)) {
  loop_vector_means[[i]] <- mean(mtcars[[i]])
}

loop_vector_means

identical(loop_vector_means, vectors_of_means)

#

lapply(mtcars, mean)
unlist(lapply(mtcars, mean))

sapply(mtcars, mean)

identical(unlist(lapply(mtcars, mean)), sapply(mtcars, mean))

#

purrr::map(mtcars, mean)
identical(purrr::map(mtcars, mean), lapply(mtcars, mean))

#

map_int(mtcars, mean)
map_lgl(mtcars, mean)
map_dbl(mtcars, mean)
map_chr(mtcars, mean)


map_dbl(ggplot2::txhousing, mean)
map_dbl(ggplot2::txhousing, mean, na.rm = TRUE)
summary(ggplot2::txhousing)

  
txhousing01 <- ggplot2::txhousing %>% 
 dplyr::select(-date, -median) %>% 
  group_by(year, month) %>%
   summarise_if(is.numeric, tibble::lst(mean, median), na.rm = TRUE) 


```


# Problem 23

Make some code that produces the following output:

```{r problem 23, echo=FALSE}

(Letters_matrix <- matrix(1:9, nrow=3, dimnames = list(letters[1:3], LETTERS[1:3])))

```

Remember to make comments. 

```{r, problem 23 solution, eval=FALSE, include=FALSE}

Letters_matrix <- matrix(1:9, nrow=3, dimnames = list(letters[1:3], LETTERS[1:3]))
Letters_matrix

```


